---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the [given names data
set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2015_txt.zip), answer some of the following questions:

- First name frequency evolves along time?
- What can we say about ``Your name here'' (for each state, all the country)?
- Is there some sort of geographical correlation with the data?
- Which state has a larger variety of names along time?
- _your own question_ (be creative)

You need to use the _dplyr_ and _magrittr_ for this analysis. Unzip the file _dpt2015_txt.zip_ (to get the **dpt2015.txt**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

```{bash}
rm -f dpt2015*.zip
wget -qc https://www.inf.ufrgs.br/~schnorr/tmp/dpt2015_txt.zip
unzip -o dpt2015_txt.zip
ls -lh dpt2015*
```

```{r}
library(readr);
df <- read_tsv("dpt2015.txt", locale = locale(encoding = "ISO-8859-1"));
df;
```

