---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the [given names data
set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2017_txt.zip), answer some of the following questions:

- First name frequency evolves along time?
- What can we say about "the french version of your name here" (for each state, all the country)?
- Is there some sort of geographical correlation with the data?
- Which state has a larger variety of names along time?
- _your own question_ (be creative)

You need to use the _dplyr_ and _magrittr_ for this analysis. Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

```{r}
file = "dpt2017_txt.zip"
if(!file.exists(file)){
  download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2017_txt.zip",
	destfile=file)
}
unzip(file)
```

```{r}
library(readr);
df <- read_tsv("dpt2017.txt", locale = locale(encoding = "ISO-8859-1"), col_types=cols(
  sexe = col_integer(),
  preusuel = col_character(),
  annais = col_character(),
  dpt = col_character(),
  nombre = col_integer()
));
df;
```

