---
title: "Estimator of the Max"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "November, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Estimation of the range of a uniform distribution
Consider a sample of random values uniformly distributed on a interval $[0,\theta]$ where $\theta$ is an unknown parameter.

The aim is to estimate the value of $\theta$.

## Initialization
```{r}
version
library(dplyr)
library(magrittr)
library(ggplot2)
```




## Statistical analysis of the estimators

### Build the generation function for a set of experiments
```{r}
library(tidyr)
Sample_Estimators <- function(Seed = 42, theta = 2, Sample_Size = 10, Experience_Number = 20) {
  set.seed(Seed)
  do.call("rbind",
    lapply(1:Experience_Number, function(x){
      y <- runif(Sample_Size, min = 0, max = theta) 
      # generates n values uniformly distributed on $[0,\theta]$
      data.frame(Experience = x, 
                 Estimator = c("Maximum", "Maximum Debiaised", "Double Average"),
                 Observation = c(max(y),  max(y)*(1+1.0/Sample_Size), 
                         mean(y)*2) 
                 )
    }
    )
  ) 
}
Sample_Estimators( Experience_Number = 10)
```

### Fix the experimental parameters for the simulation
```{r}
# Experimental parameters
theta = 2 # size of 
Sample_Size = 100
Experience_Number = 100
Seed = 42

# plot parameters
Order_plot =c("Maximum", "Maximum Debiaised", "Double Average")
```

### Plot a summary of the distribution of the estimators
```{r}

Title_graphic = paste("Statistics of estimators of the maximum\n",
                      " Theta:",theta,
                      " Sample Size per experiment : ",  Sample_Size,
                      " Number of experiments:",Experience_Number)

Sample_Estimators(Seed, theta , Sample_Size , Experience_Number ) %>% 
  mutate(Estimator = factor(Estimator,Order_plot)) %>%
  ggplot(aes(x = Estimator, y = Observation)) +
  geom_boxplot(width = 0.3, color = "darkred") +
  stat_summary(fun.y=mean, colour="black", geom="point", 
               shape=4, size=5, show.legend = FALSE) + 
  geom_hline(yintercept=2, color="blue") +
  xlab("") +
  ylab("Statistics of the estimator") +
  theme_bw(base_size=12) +
  ggtitle(Title_graphic) +
  coord_cartesian(ylim=c(1.6,2.3))
```

### Analysis  of the plot

In the figure we have an idea about the differences between the estimators. The mean of the estimators, $\times$ points, show the biais of the _Maximum_ estimator, the height of the boxplot illustrates the risk associated with each estimator. Even if the first estimator is unbiaised the associated risk seems smaller than the _Double Average_ estimator. This should be confirmed by some tests. 

### To go further 
We can plot the histograms of the distributions of the estimators

```{r}
Sample_Estimators(Seed, theta, Sample_Size=1000, Experience_Number) %>% 
  mutate(Estimator = factor(Estimator,Order_plot)) %>%
  ggplot(aes(x=Observation, y=..density..)) +
  geom_vline(xintercept=2, color="blue") +
  geom_histogram(binwidth = 0.005) + 
  coord_cartesian(xlim=c(1.5,2.5)) + 
  theme_bw() +
  facet_wrap(~Estimator, ncol=1)
  # ggsave(filename = "hist.png")
```
Exercise : why these representations are misleading ? Explain how you can avoid this misleading.

