---
title: "How to Break a Stick ?"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2017"
output: pdf_document
---


# Break a stick in two pieces
=============================

Consider a stick, break it in two pieces. How much times the larger piece is on the smaller one ?

0--------------------X----------------------1

Modelling
------------

Suppose that the length of the stick is $1$, the point of break is modelled by a random variable $X$ with value in $[0,1]$. The largest piece is $\max(X,1-X)$, the smaller one $\min(X,1-X)$.

The variable of interest $Z$ is defined by :
\[
Z=\frac {\max(X,1-X)}{\min(X,1-X)}.
\]

The aim of this activity is to study the average of $Z$ and our approach will be the simulation.

*Statistical hypothesis* : $X$ is uniformly distributed on  $[0,1]$ 

To ensure the reproducibility of our simulations we fix the seed of the random generator.
## Initialization
```{r}
version
set.seed(1:6)
library(dplyr)
library(magrittr)
library(ggplot2)

```

# Basic sample set generation

```{r}
Stick_Uniform <- function(n=10) {
  x <- runif(n) # generates n cuts uniformly distributed on $[0,1]$
  data.frame(n=c(1:n), X=x, Y=1-x) %>% 
    mutate(Z=pmax(X,Y)/pmin(X,Y)) %>%
    select(n,X,Z)
}
Stick_Uniform(100)
```
## First observations 
----------------------------------
Generate a sample of $n$ points break and compute the average

We use a single sample Sample_Stick_Uniform (with a large size) 
```{r}
  Sample_Stick_Uniform <- Stick_Uniform(100000)
```

Let's try to make the computation of the average of $Z$ with a sample of size $1000$
```{r fig.width=3, fig.height=3}
Stick_summary <- function(df,sample_size=10) {
  # compute the summary/boxplot of the first n rows
    df %>% slice(1:sample_size) %>%
    summarise(mean = mean(Z))
}

Stick_summary(Sample_Stick_Uniform,1000)
```
Try other valuse of the sample size

Compute now the partial average depending on the size of the sample



Plot the average of the value depending on the size of the sample

```{r}
Stick_Evolution <- function(df) {
  df %>%
    mutate(CS=cumsum(Z)) %>% 
    mutate(AVG=CS/n) 
}  

Stick_Evolution(Sample_Stick_Uniform)  %>% 
    ggplot(aes(x=n,y=AVG)) + geom_line() + ylim(0,200) + theme_bw()
```
For several simulations
```{r}
set.seed(43)
Multiple_Sample_Uniform <-do.call("rbind",lapply(1:10, function(x){
  Stick_Evolution(Stick_Uniform(n=100000)) %>%
    mutate(Q=x)
})) 
Multiple_Sample_Uniform %>%
  ggplot(aes(x=n,y=AVG,color=as.factor(Q))) + 
    geom_line() + 
    coord_cartesian(ylim=c(0,100)) + 
    theme_bw()
```

As we can see on the plot, the average is not converging to a same limit, and so depends deeply on the sample size. We are in a case when the variability is too high to define an average.

Moreover, we can plot an histogram 

```{r}
Sample_Stick_Uniform %>%
    ggplot(aes(Z)) + geom_histogram(bins = 200) + xlim(0,500)

```



## Probabilistic Model

What the theory says ?

We compute the probability distribution of $Z$, that is  $\mathbb{P}(Z \leq z)$.

First note that the $\min (x,1-x)$ is uniformly distributed on $[0,\frac 12]$, and $\max (x, 1-x) = 1-\min(x,1-x)$. 

\[
\mathbb{P}(Z \leq z) = \mathbb{P}\left(\frac {1-V}V \leqslant z\right) \text{ where } V\sim \mathcal{U}_{[0,\frac 12]}.
\]
\[
\mathbb{P}(Z \leq z) = \mathbb{P}\left(\frac {1-V}V \leqslant z\right) = 
\mathbb{P}\left( {1-V} \leqslant zV\right)=\mathbb{P}\left(\frac 1{1+z}\leqslant V\right) = 2 \left(\frac 12 - \frac 1{1+z}\right)= \frac {z-1}{z+1}.
\]
The density is obtained by derivation :
\[
f_Z(z) = \frac 1{(1+z)^2}
\]

And a small computation shows $\mathbb{E} Z =\int_{0}^{+\infty}z \frac 1{(1+z)^2}dz= \left[\frac 12 \log (1+z^2)\right]_0^{+\infty} = +\infty$

## Other distribution of the break point ?

For example with a triangular distribution on the same set with the same average $\frac 12$
```{r}
Stick_Triangular <- function(n=10) {
  x <- (runif(n)+runif(n))/2  # generates n cuts distributed according the triangular distribution
  data.frame(n=c(1:n), X=x, Y=1-x) %>% 
    mutate(Z=pmax(X,Y)/pmin(X,Y)) %>%
    select(n,X,Z)
}

set.seed(43)

Multiple_Sample_Triangular <- do.call("rbind",lapply(1:10, function(x){
  Stick_Evolution(Stick_Triangular(n=100000)) %>%
    mutate(Q=x)
})) 

Multiple_Sample_Triangular %>%
  ggplot(aes(x=n,y=AVG,color=as.factor(Q))) + 
    geom_line() + 
    coord_cartesian(xlim=c(0,2000),ylim=c(0,20)) + 
    theme_bw()
```

What conclusions could be drawn on this new way of cutting ?
